### 2020-03-09 short.body.tex
various tweaks
Tweaks to Discussion; nuke commented-out old Discussion for some reason

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.45\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.45\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the realized proportion of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Increasing the speed of asymptomatic transmission (shorter generation intervals) increases the realized proportion of asymptomatic transmission, $q$.
(B) Increasing the speed of asymptomatic transmission (shorter generation intervals) decreases the basic reproduction number ${\cal{R}}_0$.
When $\bar G_a$ is smaller (larger) than $\bar G_s$, estimates based on the observed generation distribution for symptomatic cases (${\cal{R}}_0=2.5$; dashed line) are expected to over- (under-) estimate the true $\mathcal R_0$.
Triangle assume $z=0.5$ and $\bar G_a/\bar G_s = 0.55$.
Circle assume $z=0.5$ and $\bar G_a/\bar G_s = 1.8$.
Solid lines show contours for \Ro\ values. 
Here, we assume $1/r=7$ days, $\bar G_s=8$ days, and $\kappa_s=\kappa_a=0.5$.
\label{fig.importance}}
\end{center}
\end{figure*}

@@

Much is still unknown about course of asymptomatic infection compared to symptomatic infection in COVID-19. 
Our current work highlights the need to characterize the duration of asymptomatic infections, relative to that of symptomatic ones, and their consequences not only for contact tracing but for precise estimation of the basic reproduction number of the ongoing COVID-19 outbreak~\citep{park_preprint}.
The present findings are also consistent with a recent generalization linking speed, strength, and generation intervals, such that for a given observed speed increases in the mean generation interval imply larger reproduction number~\citep{park_2019practical}.

If mild, asymptomatic infections are more persistent than severe, symptomatic ones, the mean generation interval for COVID-19 could be longer than estimated from symptomatic cases alone - possibly increasing ${\cal{R}}_0$ (Figure~\ref{fig.importance}B).
However, if asymptomatic cases tend to have shorter infection, then current estimates of ${\cal{R}}_0$ may be over-estimates of the underlying strength (Figure~\ref{fig.importance}B), although asymptomatic cases may be driving a larger fraction of secondary cases than we would expect without accounting for their differences (Figure~\ref{fig.importance}A).
Therefore, current estimates of ${\cal{R}}_0$ that do not account for asymptomatic cases may be systematically biased.

Finally, it is important to note we focused here on the exponential phase, but that the realized proportion of asymptomatic transmission $q$ is time-dependent, and will vary given the population state structure of infected individuals. Such variation in the ratio of asymptomatic to symptomatic cases will also impact estimates of case fatality rates insofar as many asymptomatic cases are not counted, particularly early in an outbreak.
Future work might also consider the ways in which asymptomatic individuals modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.
Characterizing the role of asymptomatic individuals in driving the persistence of the epidemic will be critical for assessing the post-pandemic outcome \citep{lipsitch_preprint}.

%We also note that efforts to characterize asymptomatic tranissmion
%should consider its effect on
%the basic reproduction number, which for this SEIR coronavirus-like model is ${\cal{R}}_0=p\left(\frac{\beta_a}{\gamma_a}\right)+ (1-p)\left(\frac{\beta_s}{\gamma_s}\right).$
%%Pooling estimates via a Bayesian approach applied to the early stages of Wuhan case data,
%%suggests ${\cal{R}}_0\approx 3.0$ 
%%with a 95\% confidence interval from 2-4.5.  Yet, that information
%%is not sufficient to disentangle the role of asymptomatic cases, i.e.,
%%what fraction of secondary cases can be ascribed to transmission
%%from asymptomatic cases vs.~symptomatic cases?  
%Strength, ${\cal{R}}_0$ and size $r$
%can be related via a Gamma approximation to the generation
%interval distribution as follows:
%\begin{equation}
%{\cal{R}}^{gamma}=(1+\kappa r \bar{G})^{1/\kappa}
%\end{equation}
%where $\bar{G}$ is the mean generation interval and
%$\kappa$  is the squared coefficient of variation of
%the generation interval.  
%This approximation (explored in Champredon
%et al.) reveals why for a given observed speed,
%increases in the mean generation interval imply
%larger reproduction number. Hence, reports of individuals
%with mild cases that remain infectious for (relatively)
%long periods suggest that the mean generation interval
%could be longer than estimated from symptomatic cases alone - both
%increasing ${\cal{R}}_0$ and driving a larger fraction of secondary cases.
%This inference also reinforces 
%the need to characterize the epidemiological characteristics of asymptomatic transmission.

%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and $M(r)$ is the
%moment generating function of the generation interval distribution
%$g(a)$ (normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. The relevance $q$ is an unknown feature of
%outbreaks at the population scale, distinct from $p$ the probability of having
%an asymptomatic cases, conditional upon being infected
%but not yet infectious. Hence,
%if one were to observe $r$ (which can vary from country to 
%country and from region to region) then it should be possible
%to link the values $p$ and the relative transmission rates,
%$\beta_a/\beta_s$ to the relevance of asymptomatic transmission
%at the population scale.

%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and $M(r)$ is the
%moment generating function of the generation interval distribution
%$g(a)$ (normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. The relevance $q$ is an unknown feature of
%outbreaks at the population scale, distinct from $p$ the probability of having
%an asymptomatic cases, conditional upon being infected
%but not yet infectious. Hence,
%if one were to observe $r$ (which can vary from country to 
%country and from region to region) then it should be possible
%to link the values $p$ and the relative transmission rates,
%$\beta_a/\beta_s$ to the relevance of asymptomatic transmission
%at the population scale.
%
%
%
%
%Hence, recalling the equation for the basic reproduction number
%\begin{equation}
%{\cal{R}}_{0,E}= p\left(\frac{\beta_a}{\gamma_a}\right)+ (1-p)\left(\frac{\beta_s}{\gamma_s}\right)
%\end{equation}
%suggests that 
%This gap between intrinsic and realized transmission routes
%lies at the base of our exploration.
%
%\section{Dynamics}
%We simulate the SEIaIsRD model assuming $N=10^6$ with
%the following initial conditions: $S_0=10^6-1$, $E_0=1$
%and all other elements 0.  The following model vary $p$ from
%0.1 to 0.5 to 0.9.  For various reasons, the parameters are:
%$\gamma_e=1/4$, $\gamma_a=1/42$, $\gamma_s=1/14$,
%$\beta_a=1/14$, $\beta_s=3/14$, $f=0.02$.  The resultant
%value of ${\cal{R}}_0=3$ for all values of $p$, however
%the realized CFR, fraction of asymptomatic cases, and even
%the speed of the disease varies significantly. Hence, we 
%need to adopt a different approach, given that we
%observe a speed, $r$, and yet are uncertain with
%respect to the underlying contribution of asymptomatic
%vs.~symptomatic cases.
%
%\begin{figure*}[h!]
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figseird_base_noname.pdf}\\
%\includegraphics[width=0.9\textwidth]{figseird_base2_noname.pdf}\\
%\includegraphics[width=0.9\textwidth]{figseird_base3_noname.pdf}\\
%\end{center}
%\end{figure*}
%
%\section{Constrained Dynamics}
%Here, we simulate the SEIR model developed
%for coronavirus assuming $N=10^6$ with
%the following initial conditions: $S_0=10^6-1$, $E_0=1$
%and all other elements set to 0.  We assume that the speed of
%the outbreak is $r=1/6$ days$^{-1}$. We further assume that $\gamma_e=1/4$
%(4 days to onset), $\gamma_a=1/42$, $\gamma_s=1/14$, and $p=0.9$.
%How do we solve for the unknown transmission rates, $\beta_a$
%and $\beta_s$?  Alone, it would seem we have a under-determined
%problem, such that
%\begin{equation}
%p\left(\frac{\beta_a}{\gamma_a}\right)+ (1-p)\left(\frac{\beta_s}{\gamma_s}\right)=\frac{1}{M(-r)}.
%\end{equation}
%However, if one assumes that a fraction $q$ of secondary infections are caused by
%asymptomatic individuals, then there is an additional constraint:
%\begin{equation}
%p\frac{\beta_a}{\gamma_a}=\frac{q}{M(-r)}
%\end{equation}
%such that 
%\begin{equation}
%\beta_a = \frac{q\gamma_a}{pM(-r)}.
%\end{equation}
%As a result, then 
%\begin{equation}
%\beta_s = \frac{(1-q)\gamma_s}{(1-p)M(-r)}
%\end{equation}
%In essence, as $q$ varies, then both tranmission rates will have 
%to covary (indeed in a negative fashion) to reconciled the realized
%speed of the outbreak. We now put this into practice for three scenarios,
%all for the case of $p=0.9$ (in which intrinsically most cases
%are asymptomatic):
%\begin{description}
%\item[Short asymptomatic cases] Here we assume
%$\gamma_s=1/14$ (2 week for severe cases) and $\gamma_a=1/2$ 
%(2 day resolution, i.e., a 7-fold change). 
%\item[Long asymptomatic cases] Here we assume
%$\gamma_s=1/14$ (2 week for severe cases) and $\gamma_a=1/42$ 
%(6 week for mild cases)
%\end{description}
%We also assume as before that $f=0.02$ for severe cases. The following 
%results show that we can in fact calculate the correct values of
%$\beta_a$ and $\beta_s$ given variation in $q$.
%It is also important to point out that
%${\cal{R}}_0$ can be extended or decreased with $q$ depending
%on whether $\gamma_a$ is faster or slower relative to $\gamma_s$.
%The key idea there is the same as that for the impact of post-death
%transmission of Ebola virus disease.
%
%For the short duration asymptomatic cases (see Figure~\ref{constrain}), we find the following trade-offs:
%\begin{verbatim}
%         q: [0.0100 0.1000 0.5000 0.9000 0.9900]
%    beta_a: [0.0258 0.2311 0.7857 1.0714 1.1176]
%    beta_s: [3.2898 2.6741 1.0102 0.1531 0.0145]
%        R0: [4.6523 4.1597 2.8286 2.1429 2.0320]
%\end{verbatim}
%whereas for the longer duration asymptomatic cases we find
%\begin{verbatim}
%         q: [0.0100 0.1000 0.5000 0.9000 0.9900]
%    beta_a: [0.0013 0.0132 0.0873 0.2311 0.2843]
%    beta_s: [3.3528 3.2143 2.3571 0.6933 0.0775]
%        R0: [4.7414 5 6.6000 9.7059 10.8553]
%\end{verbatim}
%In the first case, when most illnesses
%are caused by asymptomatic cases, then shorter generation intervals
%for the same growth rate means a lower ${\cal{R}}_0$. Whereas,
%in the second case, when most illnesses are caused by asymptomatic
%cases, then longer generation intervals for the same growth rate
%means a higher ${\cal{R}}_0$.
%\begin{figure*}[b!]
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figconstrain_short_noname.pdf}\\
%\includegraphics[width=0.9\textwidth]{figconstrain_long_noname.pdf}\\
%\caption{Short, mild asymptomatic cases (top) vs. longer, mild asymptomatic
%cases (bottom). For both, common parameters are $\gamma_e=0.25$,
%$\gamma_s=1/14$, $f=0.02$, $p=0.9$, and $r=1/7$.  For the short
%asymptomatic case scenario, $\gamma_a=1/2$ whereas for the long asymptomatic
%case scenario $\gamma_a=1/42$.
%\label{constrain}}
%\end{center}
%\end{figure*}
%
%Finally, in an effort to try and reconcile this, we point out the following
%synthetic figure of outbreaks all with a growth rate of $r=1/7$ by construction, yet differ in vastly different ways based on 3 facets, the value of $p$ (the instrinsic fraction of asymptomatic cases),
%the value of $1/\gamma_a$ (that sets the average duration of asymptomatic cases), and
%the realized value of $q$ (the fraction of secondary cases caused by asymptomatic cases).  As is
%apparent, all disease take off at the same speed, but have very different peak strengths,
%and also, different decomposition of the relative importance of the asymptomatic component.
%What is notable is that the peak size is more driven by $p$ and $\gamma_a$ than by $q$
%(see Figure~\ref{constrain2}).
%\begin{figure*}[bh!]
%\begin{center}
%\includegraphics[width=0.9\textwidth]{package_overlay_p_T.pdf}
%\caption{Synthesis of dynamical results given variation in 
%the intrinsic generation interval of asymptomatic cases (set by $1/\gamma_a$), frequency ($p$), 
%and importance ($q$).
%For all, common parameters are $\gamma_e=0.25$,
%$\gamma_s=1/14$, $f=0.02$, and $r=1/7$.  For the short
%asymptomatic case scenario, $\gamma_a=1/2$ whereas for the long asymptomatic
%case scenario $\gamma_a=1/42$.
%\label{constrain2}}
%\end{center}
%\end{figure*}
%
%%
%%\section{Simulations}
%%Consider an outbreak whose
%%symptomatic cases is ${\cal{R}}_0\approx 3$, such that
%%$\beta_s/\gamma_s=3$.  We further assume that transition from
%%exposed to either $I_a$ or $I_s$ takes 4 days on average,
%%such that $\eta=1/4$.  The following is an example
%%outbreak, in which $\beta_s=\beta_a=3/7$, $\gamma_a=1/45$,
%%and $f=0.02$ (corresponding to a 2\% case fatality rate
%%for symptomatic cases). As is apparent, there is an exponential
%%increase in cases, such that $r\approx 0.30$ days$^{-1}$.
%%In this case, we expect that in the exponential phase,
%%then $I_a/I_s\approx 13.8$.  This prediction is borne out
%%by simulation as in Figure~\ref{fig.ratio}.
%%\begin{figure*}[h!]
%%\begin{center}
%%\includegraphics[width=0.4\textwidth]{figexample1_noname.pdf}
%%\mbox{\hspace{0.05\textwidth}}
%%\includegraphics[width=0.4\textwidth]{figexample1_ratio_noname.pdf}
%%\caption{Asymptomatic to symptomatic case ratio at 
%%the population scale can exceed the ratio of relative case
%%production at the individual scale, when $\gamma_a\ll \gamma_s$.\label{fig.ratio}}
%%\end{center}
%%\end{figure*}
%%Similalry, in the case that asymmptomatic cases rapidly
%%recover, then we expect the population rate to be less then
%%$z$. In the event that $\gamma_a=1$ (7-fold faster than
%%the the symptomatic removal rate) then $r\approx 0.17$ and
%%the ratio is expected to converge to $\approx 2.7$, as shown in Figure~\ref{fig.ratio2}.
%%\begin{figure*}[h!]
%%\begin{center}
%%\includegraphics[width=0.4\textwidth]{figexample2_noname.pdf}
%%\mbox{\hspace{0.05\textwidth}}
%%\includegraphics[width=0.4\textwidth]{figexample2_ratio_noname.pdf}
%%\caption{Asymptomatic to symptomatic case ratio at 
%%the population scale can exceed the ratio of relative case
%%production at the individual scale, when $\gamma_a\gg \gamma_s$.
%%\label{fig.ratio2}
%%}
%%\end{center}
%%\end{figure*}
%%Finally, in the case that asymptomatic cases and symptomatic
%%cases have the same exepcted rudation
%%then we expect the population ratio to be less then
%%$z$. In the event that $\gamma_a=1$ (7-fold faster than
%%the the symptomatic removal rate) then $r\approx 0.26$ and
%%the ratio is expected to converge to $\approx 10$, as shown in Figure~\ref{fig.ratio3}.
%%\begin{figure*}[h!]
%%\begin{center}
%%\includegraphics[width=0.4\textwidth]{figexample3_noname.pdf}
%%\mbox{\hspace{0.05\textwidth}}
%%\includegraphics[width=0.4\textwidth]{figexample3_ratio_noname.pdf}
%%\caption{Asymptomatic to symptomatic case ratio at 
%%the population scale can exceed the ratio of relative case
%%production at the individual scale, when $\gamma_s=\gamma_a$.
%%\label{fig.ratio3}
%%}
%%\end{center}
%%\end{figure*}
%%
%%\section{Stability of the infection-free equilibrium}
%%Consider the following infection subystem, assume $S\approx 1$
%%\begin{eqnarray}
%%\dot{E}&=&\beta_a I_a +\beta_s I_s -\eta E\\
%%\dot{I}_a&=&\frac{z}{z+1}\eta E-\gamma_a I_a\\
%%\dot{I}_s&=&\frac{1}{z+1}\eta E-\gamma_s I_s
%%\end{eqnarray}
%%such that the Jacobian is:
%%\begin{equation}
%%J = \left[
%%\begin{array}{ccc}
%%-\eta & \beta_a & \beta_s \\
%%\frac{z}{z+1}E & -\gamma_a & 0 \\
%%\frac{1}{z+1}E & -\gamma_s & 0 
%%\end{array}
%%\right]
%%\end{equation}

>>>>

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.45\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.45\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the realized proportion of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Increasing the speed of asymptomatic transmission (shorter generation intervals) increases the realized proportion of asymptomatic transmission, $q$.
(B) Increasing the speed of asymptomatic transmission (shorter generation intervals) decreases the basic reproduction number ${\cal{R}}_0$.
When $\bar G_a$ is smaller (larger) than $\bar G_s$, estimates based on the observed generation distribution for symptomatic cases (${\cal{R}}_0=2.5$; dashed line) are expected to over- (under-) estimate the true $\mathcal R_0$.
Triangle assume $z=0.5$ and $\bar G_a/\bar G_s = 0.55$.
Circle assume $z=0.5$ and $\bar G_a/\bar G_s = 1.8$.
Solid lines show contours for $q$ and \Ro\ values. 
Here, we assume $1/r=7$ days, $\bar G_s=8$ days, and $\kappa_s=\kappa_a=0.5$.
\label{fig.importance}}
\end{center}
\end{figure*}

@@

Much is still unknown about the time scale and effectiveness of asymptomatic transmission in COVID-19. 
Here we highlight the need to characterize the generation-interval distribution for asymptomatic transmission, and its consequences not only for contact tracing but for estimation of the basic reproduction number of the ongoing COVID-19 outbreak~\citep{park_preprint} and of the effective proportion of asymptomatic transmission during the exponential-growth phase.
Our reproductive number findings fit into a broader framework linking epidemc speed, strength, and generation intervals -- for a given observed speed increases in the mean generation interval imply larger reproduction number~\citep{park_2019practical}. \jd{Add one or two other cites, maybe starting with Wallinga Lipsitch.}

If asymptomatic infections are more persistent than symptomatic ones, the mean generation interval for COVID-19 could be longer than estimated from symptomatic cases alone -- possibly increasing ${\cal{R}}_0$ (Figure~\ref{fig.importance}B).
However, if asymptomatic cases tend to resolve quickly, then current estimates of ${\cal{R}}_0$ may be over-estimates of the underlying strength (Figure~\ref{fig.importance}B), and asymptomatic cases may be driving a larger fraction of secondary cases than we would expect without accounting for their differences (Figure~\ref{fig.importance}A).

We focus here on the exponential phase, so it is worth noting that the realized proportion of asymptomatic transmission $q$ is time-dependent, varying with dynamic changes in incidence and proportion susceptible. 
Future work might also consider the ways in which asymptomatic individuals can modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.
Characterizing the role of asymptomatic individuals in driving the persistence of the epidemic will be critical for assessing the post-pandemic outcome \citep{lipsitch_preprint}.
